### 2021-05-06 submission/manuscript.Rmd
Add to discussion

## Discussion
By following osmotic laxative treated mice over time, we found 5-day and 1-day PEG treatments exhibited either prolonged or transient *C. difficile* colonization after challenge. The differences in *C. difficile* colonization dynamics between the 5- and 1-day PEG treatments were associated with differences in how much the treatments disrupted the microbiota. We also found the intestinal communties of 5-day PEG treated mice had not regained colonization resistance after a 10-day recovery period. Additionally, treating mice with PEG after *C. difficile* challenge prolongs colonization instead of enhancing clearance, as suggested by the literature. Overall, our results demonstrated that treating mice with an osmotic laxative rendered mice susceptible to *C. difficile* colonization and duration of colonization depended on the length of PEG treatment and whether treatment was administered before or after *C. difficile* challenge. Note (remove figure 8?)

* Discussion of prolonged persistence. *C. difficile* sequences detected in tissue samples. Association with mucin-degrading bacteria suggested by recent papers.

* Discuss why we might not have observed more severe histology in PEG mice relative to clindamycin-treated mice
  + Antibiotics may also impact mucus layer
  + Strain of bacteria used

@@

* Discuss what these findings might mean for human patients (Fig. 8B)
  + What's known regarding laxatives and susceptibility to CDIs
  + Clinical trial of PEG, results never posted [@Dieterle2018]
  + Relevance to human FMTs? Unclear what the best administration route is because there have been no studies designed to evaluate the best administration route for FMTs.

>>>>

## Discussion
By following osmotic laxative treated mice over time, we found 5-day and 1-day PEG treatments exhibited either prolonged or transient *C. difficile* colonization after challenge. The differences in *C. difficile* colonization dynamics between the 5- and 1-day PEG treatments were associated with differences in how much the treatments disrupted the microbiota. The intestinal communities of 5-day PEG treated mice had not regained colonization resistance after a 10-day recovery period. In contrast to the hypothesize suggested by the literature, treating mice with PEG after *C. difficile* challenge prolonged colonization. We identified patterns in the relative abundances of *Bacteroides*, *Enterobacteriaceae*, *Akkermansia*, *Porphyromonadaceae*, and *Lachnospiraceae* that were associated with prolonged *C. difficile* colonization (Fig. 8). Overall, our results demonstrated that osmotic laxative treatment alone rendered mice susceptible to *C. difficile* colonization and the duration of colonization depended on the length of PEG treatment and whether treatment was administered before or after *C. difficile* challenge. Note (remove figure 8B, check 8A...specify 5-day vs 1-day or just  in figure legend)

Some of the lower abundance bacteria could also play important roles in clearing C. difficile. For example, the relative abundance of Oscillibacter was increased in mice that cleared *C. difficile* compared to the mice with prolonged colonization. Oscillibacter __ can produce the SCFA valerate, which has been shown to inhibit *C. difficile* growth based on ___ data. Similarly, Acetatifactor can produce __, another SCFA with ___ role in the context of CDIs. Thus bacteria with potential protective functions might have played a role in promoting *C. difficile* clearance.

There was an interesting pattern of increased Enterobacteriaceae, Bacteroides, and C. difficile in both the stool and mucosal communities of PEG-treated mice suggesting a potential synergy. 
Akkermansia is also a mucin degrader with potentially beneficial or detetrimental roles depending on context in other diseases (). In our study the relative abundance of *Akkermansia* shifted over time between groups. In the stool it was initially increased in mice that cleared *C. difficile*, but shifted after 5-days post-challenge so that it was increased in mice that had prolonged colonization. In the context of CDIs, some studies suggest a protective role, while others suggest detrimental.

* Discussion of prolonged persistence. *C. difficile* sequences detected in tissue samples. Association with mucin-degrading bacteria suggested by recent papers. Akkermansia, Bacteroides are known mucin degraders and could be compete.

Bacteroides and Enterobacteriaceae synergy in the context of CRC (Dejea paper?)

@@

Despite identifying an altered compositional profile that included higher relative abundance of the *C. difficile* sequence in the mucosal tissues of mice treated with 5-day PEG compared to the clindamycin group, we did not see a difference in histopathology scores between the groups. One reason there was no difference could be the *C. difficile* strain used, *C. difficile* 630 is mild in mice compared to VPI (ref). Part of our hypothesis for why we might see increased histopathology scores is because PEG was previously shown to disrupt the mucus layer in mice. However, a recent study demonstrated that broad spectrum antibiotics (clinda included) can also disrupt the host mucus layer in mice? ref. Birchenough. Future research is needed to tease out the interplay between drugs that influence the mucus layer and different strains of *C. difficile* in the context of CDIs.

Discuss why we might not have observed more severe histology in PEG mice relative to clindamycin-treated mice
  + Antibiotics may also impact mucus layer
  + Strain of bacteria used

* Discuss what these findings might mean for human patients 
  + What's known regarding laxatives and susceptibility to CDIs
  + Clinical trial of PEG, results never posted [@Dieterle2018]
  + Relevance to human FMTs? Unclear what the best administration route is because there have been no studies designed to evaluate the best administration route for FMTs.
